---
title: "Stock investment analysis"
output: html_document
date: "2025-02-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 2
## This is an investment dataset from 2015-2018, provides 25 variables with more than 100,000 observations
```{r cars}
library(tidyverse)
investment <- read_csv("investment_transactions.csv")
head(investment)
```
## 3
```{r}
str(investment)
```
## 4
```{r}
summary(investment)
```
### 5
Binary: investment
Nominal (Categorical, No Order): company, sector
Ordinal (Categorical, Has Order): ESG_ranking 
Discrete: horizon (days), inflation
Continuous: price_BUY, price_SELL, PE_ratio, EPS_ratio

## 6
1. Is there a correlation between ESG ranking and return on equity?
2. How does the volatility of an asset impact its sharpe ratio?
3. Do investments in the bank sector have higher expected returns than those in retail?

## 7
## Analyze the relationship between buy and sell price
```{r}
ggplot(investment, aes(x = price_BUY, y = price_SELL, colour = sector)) +
  geom_point() +
  geom_smooth(method="lm", color = "brown") +
  labs(
       x = "Buy Price",
       y = "Sell Price")
```

## 8
## Measure how buy volatility ratio affect the amount of return
```{r}
investment <- investment %>%
  mutate(return_volatility_ratio = nominal_return / Volatility_Buy)
head(investment)
```

## 9
## Compare 2 sectors in term of return on assets ratio
```{r}
investment %>%
  filter(sector %in% c("BANK", "TECH")) %>%
  ggplot(aes(x = company, y = roa_ratio, colour = sector)) +
  geom_boxplot() +
  labs(title = "",
       x = "Company",
       y = "Return on Assets Ratio")
```

## Pearson correlation between buy and sell price
```{r}
cor.test(investment$price_BUY, investment$price_SELL)
```

## Evaluate the net profit percentage of highest-trading-volume sector with lowest one to find out if company's profit lead to buy/sell decision
```{r}
filtered_company <- investment %>%
  filter(sector %in% c("RETAIL", "FMCG"))

# Create the boxplot
ggplot(filtered_company, aes(x = company, y = NetProfitMargin_ratio, colour = sector)) +
  geom_boxplot() +
  scale_y_continuous(
    breaks = seq(-20, 60, by = 5)
  ) +
  labs(title = "Net Profit Ratio of companies in FMCG and Retail industries",
       x = "Company",
       y = "Net Profit Ratio")

lr_amount <- lm(amount ~ NetProfitMargin_ratio, data = investment)
summary(lr_amount)
```

## Detect which sectors has highest trading volume and which has lowest
```{r}
result <- investment %>%
  group_by(sector) %>%
  summarize(total_amount = sum(amount))
```

## Find out which pair of variables has probably high relationship and analyze those 2 variables
```{r}
cor(select(investment, Volatility_Buy, Volatility_sell, 
    `Sharpe Ratio`, `expected_return (yearly)`, inflation, 
    nominal_return, ESG_ranking, PE_ratio, EPS_ratio, 
    PS_ratio, PB_ratio, NetProfitMargin_ratio, 
    current_ratio, roa_ratio, roe_ratio))

investment %>%
  filter(roe_ratio > 0) %>%
  ggplot(aes(x = roe_ratio, y = roa_ratio)) +
  geom_point() +
  geom_smooth(method="lm") +
  labs(
       x = "Return on Equity",
       y = "Return on Assets")

lr <- lm(EPS_ratio ~ NetProfitMargin_ratio, data = investment)
summary(lr)
```

## In conclusion, the variables in investment dataset is great for evaluating the development and performance of companies as well as industries. 
## However, while these variables contribute to investors' decision-making, they don't provide a perfect linear relationship of trading volume, as there are factors besides companies that investors need to consider.

